The \dealii{} Library, Version 9.3

### 2021-04-23 9.3/paper.tex
Change affiliation for marcfehling

 \author[2]{Wolfgang~Bangerth}
 \affil[2]{Department of Mathematics and Department of Geosciences, Colorado State University, Fort
   Collins, CO 80523-1874, USA.
   \texttt{bangerth@colostate.edu}}

@@

% \author[4]{Thomas~C.~Clevenger}
% \affil[4]{School of Mathematical and Statistical Sciences,
%   Clemson University,
%   Clemson, SC, 29634, USA
%   {\texttt{tcleven/heister@clemson.edu}}}
%
% \author[4]{Denis~Davydov}
% \affil[4]{Chair of Applied Mechanics,
%   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
%   Egerlandstr.\ 5,
%   91058 Erlangen, Germany.
%   {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
%
\author[5]{Marc~Fehling}
\affil[5]{Institute for Advanced Simulation,
  Forschungszentrum J\"{u}lich GmbH,
  52425 J\"{u}lich, Germany.
  {\texttt{m.fehling@fz-juelich.de}}}
%
% \author[6]{Daniel Garcia-Sanchez}
% \affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
%   Institut des NanoSciences de Paris, F-75005, Paris, France
%   {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
%
% \author[2]{Graham Harper}
%

@@

\subsubsection{Generating meshes }

@@

\subsubsection{Mixed mesh}
For mixed meshes, concepts known from the hp context have been applied: 
 different finite-element classes are assigned to different cell types. In the 
case of a 2D mixed mesh, which can only consists of triangles and
quadrilaterals, the finite element defined on a triangle (e.g., \texttt{FE\_SimplexP})
and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection}:

@@

The introduction of simplex and mixed meshes leads to some implications
for the user if these features should be used. For instance, each cell might have a different type with 
different number of vertices, lines, and faces so that these quantities can not be 
compile-time constants anymore. This information used to be queried from
the \texttt{GeometryInfo} class. To avoid using this class, we have extended
relevant classes, e.g., \texttt{TriaAccessor} or \texttt{TriaCellAccessor},
with useful new functions like \texttt{n\_vertices()}, \texttt{n\_lines()}, or
\texttt{n\_faces()} so that users can simply write:
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(unsigned int f = 0; f < cell->n_faces(); ++f)
    (void) cell->face(f);
\end{c++}
Alternatively, one can use an iterator-based approach to loop over all faces
of a cell, introduced in the last release. The relevant functions have been adjusted to be able to deal with
various number of faces. 
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(const auto & face : cell->face_iterators())
    (void) face;
\end{c++}
Furthermore, the number of degrees of freedom might differ between cells so that
local data storage units needed for assembly might need to be resized for each 
cell as it is already good practice in the hp context:
\begin{c++}
std::vector<double> local_rhs;
for(const auto & cell : tria.active_cell_iterators())
 {
   hp_fe_values.reinit(cell);
   local_rhs.resize(hp_fe_values.get_present_fe_values().n_dofs_per_cell());
 }
\end{c++}

@@

The usage of the new transfer operators (and of some of the utility functions) in the context of a hybrid multigrid algorithm (hp-MG with AMG as coarse-grid solver) for hp problems is demonstrated in the new tutorial \texttt{step-75} (see also Subsection~\ref{subsec:steps}).

@@

The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} are highly
templated to reach high performance. In particular, the template parameters include
the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$.
For application codes that rely on operators of many different degrees (e.g., because
they use p-multigrid or hp algorithms), the instantiation
means a long compilation time.

@@

\subsubsection{Parallel matrix-free hp implementation}\label{subsubsection:mf:hp}

With release 9.1, large parts of the hp algorithms in \dealii were parallelized so that
parallel matrix-based simulations can be performed with the hp infrastructure. In the present
release, we have extended the setup routines of \texttt{MatrixFree} so that it now also
provides parallel hp support.

Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active hp FE and quadrature index and users had 
to use the function \texttt{MatrixFree::create\_cell\_\allowbreak subrange\_\allowbreak hp()} or
\texttt{::create\_cell\_\allowbreak subrange\_\allowbreak hp\_\allowbreak by\_index()} to create subranges of cells with 
the same polynomial degree. This led to user codes that were hard to read due to complicated jump tables.

The creation of subranges is now performed internally, and the non-templated versions
of the \texttt{FEEvaluation} classes are extended for the hp case. To nevertheless determine
the active hp FE and quadrature index, the current cell/face range has to be provided
the constructors of the \texttt{FEEvaluation} classes, from which the relevant information
can be deduced (in the simplex case also the face type). These changes enable
users to write matrix-free code independently of whether hp is used or not.

The new tutorial \texttt{step-75} presents how to use the new hp-related features in \texttt{MatrixFree} 
in the context of a hybrid-multigrid solver.

@@

In addition, there are a number of new tutorial programs:
\begin{itemize}
\item \texttt{step-19} TODO
\item \texttt{step-68} TODO
\item \texttt{step-75} TODO
\item \texttt{step-76} is an explicit time integrator for the
        compressible Euler equations discretized with a high-order discontinuous
        Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step67} does.
        The tutorial presents advanced topics, like the usage of cell-centric loops and
        the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} tor each high
        throughput. Furthermore, the utilization of the template parameter
        \texttt{VectorizedArrayType} and the application of lambdas to capture cell and face
        integrals are discussed.
\end{itemize}

@@

The 9.3 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_2_0_and_9_3_0.html}
{around 45 incompatible changes}; see \cite{changes93}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions; and the majority change internal
interfaces that are not usually used in external
applications. That said, the following are worth mentioning since they
may have been more widely used:
\begin{itemize}
  \item 
\end{itemize}

@@

The original \dealii{} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{multicols}{2}
  \vspace*{-36pt}
  \begin{itemize}
    \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019};
    \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
    \item For $hp$~adaptivity: \cite{BangerthKayserHerold2007};
    \item For partition-of-unity (PUM) and enrichment methods of the
          finite element space: \cite{Davydov2016};
    \item For matrix-free and fast assembly techniques:
          \cite{KronbichlerKormann2012,KronbichlerKormann2019};
    \item For computations on lower-dimensional manifolds:
          \cite{DeSimoneHeltaiManigrasso2009};
    \item For curved geometry representations and manifolds:
          \cite{HeltaiBangerthKronbichlerMola2019};
          \vfill\null  \columnbreak
    \item For integration with CAD files and tools:
          \cite{HeltaiMola2015};
    \item For boundary element computations:
          \cite{GiulianiMolaHeltai-2018-a};
    \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
          \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b};
    \item For uses of the \texttt{WorkStream} interface:
          \cite{TKB16};
    \item For uses of the \texttt{ParameterAcceptor} concept, the
          \texttt{MeshWorker::ScratchData} base class, and the
          \texttt{ParsedConvergenceTable} class:
          \cite{SartoriGiulianiBardelloni-2018-a};
    \item For uses of the particle functionality in \dealii{}:
          \cite{GLHPB18}.
          \vfill\null
  \end{itemize}
\end{multicols}

>>>>

 \author[2]{Wolfgang~Bangerth}
 \affil[2]{Department of Mathematics and Department of Geosciences, Colorado State University, Fort
   Collins, CO 80523-1874, USA.
   \texttt{bangerth/marc.fehling@colostate.edu}}

@@

% \author[4]{Thomas~C.~Clevenger}
% \affil[4]{School of Mathematical and Statistical Sciences,
%   Clemson University,
%   Clemson, SC, 29634, USA
%   {\texttt{tcleven/heister@clemson.edu}}}
%
% \author[4]{Denis~Davydov}
% \affil[4]{Chair of Applied Mechanics,
%   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
%   Egerlandstr.\ 5,
%   91058 Erlangen, Germany.
%   {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
%
\author[2]{Marc~Fehling}

%
% \author[6]{Daniel Garcia-Sanchez}
% \affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
%   Institut des NanoSciences de Paris, F-75005, Paris, France
%   {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
%
% \author[2]{Graham Harper}
%

@@

\subsubsection{Generating meshes}

@@

\subsubsection{Mixed mesh}
For mixed meshes, concepts known from the $hp$-context have been applied:
 different finite-element classes are assigned to different cell types. In the
case of a 2D mixed mesh, which can only consists of triangles and
quadrilaterals, the finite element defined on a triangle (e.g., \texttt{FE\_SimplexP})
and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection}:

@@

The introduction of simplex and mixed meshes leads to some implications
for the user if these features should be used. For instance, each cell might have a different type with
different number of vertices, lines, and faces so that these quantities can not be
compile-time constants anymore. This information used to be queried from
the \texttt{GeometryInfo} class. To avoid using this class, we have extended
relevant classes, e.g., \texttt{TriaAccessor} or \texttt{TriaCellAccessor},
with useful new functions like \texttt{n\_vertices()}, \texttt{n\_lines()}, or
\texttt{n\_faces()} so that users can simply write:
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(unsigned int f = 0; f < cell->n_faces(); ++f)
    (void) cell->face(f);
\end{c++}
Alternatively, one can use an iterator-based approach to loop over all faces
of a cell, introduced in the last release. The relevant functions have been adjusted to be able to deal with
various number of faces.
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(const auto & face : cell->face_iterators())
    (void) face;
\end{c++}
Furthermore, the number of degrees of freedom might differ between cells so that
local data storage units needed for assembly might need to be resized for each
cell as it is already good practice in the $hp$-context:
\begin{c++}
std::vector<double> local_rhs;
for(const auto & cell : tria.active_cell_iterators())
 {
   hp_fe_values.reinit(cell);
   local_rhs.resize(hp_fe_values.get_present_fe_values().n_dofs_per_cell());
 }
\end{c++}

@@

The usage of the new transfer operators (and of some of the utility functions) in the context of a hybrid multigrid algorithm ($hp$-MG with AMG as coarse-grid solver) for $hp$-problems is demonstrated in the new tutorial \texttt{step-75} (see also Subsection~\ref{subsec:steps}).

@@

The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} are highly
templated to reach high performance. In particular, the template parameters include
the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$.
For application codes that rely on operators of many different degrees (e.g., because
they use $p$-multigrid or $hp$-algorithms), the instantiation
means a long compilation time.

@@

\subsubsection{Parallel matrix-free $hp$-implementation}\label{subsubsection:mf:hp}

With release 9.1, large parts of the $hp$-algorithms in \dealii were parallelized so that
parallel matrix-based simulations can be performed with the $hp$-infrastructure. In the present
release, we have extended the setup routines of \texttt{MatrixFree} so that it now also
provides parallel $hp$-support.

Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active FE and quadrature index and users had
to use the function \texttt{MatrixFree::create\_cell\_\allowbreak subrange\_\allowbreak hp()} or
\texttt{::create\_cell\_\allowbreak subrange\_\allowbreak hp\_\allowbreak by\_index()} to create subranges of cells with
the same polynomial degree. This led to user codes that were hard to read due to complicated jump tables.

The creation of subranges is now performed internally, and the non-templated versions
of the \texttt{FEEvaluation} classes are extended for the $hp$-case. To nevertheless determine
the active FE and quadrature index, the current cell/face range has to be provided
the constructors of the \texttt{FEEvaluation} classes, from which the relevant information
can be deduced (in the simplex case also the face type). These changes enable
users to write matrix-free code independently of whether $hp$-capabilities are used or not.

The new tutorial \texttt{step-75} presents how to use the new $hp$-related features in \texttt{MatrixFree}
in the context of a hybrid-multigrid solver.

@@

In addition, there are a number of new tutorial programs:
\begin{itemize}
\item \texttt{step-19} TODO

\item \texttt{step-68} TODO

\item \texttt{step-75} demonstrates a state-of-the-art way of solving a simple
      Laplace problem using $hp$-adaptation and hybrid multigrid methods on machines
      with distributed memory. This tutorial points out particularities in porting
      serial $hp$-adaptive code for parallelization. Furthermore, it guides through
      the process of writing an efficient preconditioner for $hp$-adaptive applications.

\item \texttt{step-76} is an explicit time integrator for the
      compressible Euler equations discretized with a high-order discontinuous
      Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step-67} does.
      The tutorial presents advanced topics, like the usage of cell-centric loops and
      the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} tor each high
      throughput. Furthermore, the utilization of the template parameter
      \texttt{VectorizedArrayType} and the application of lambdas to capture cell and face
      integrals are discussed.

\item \texttt{step-77} TODO

\item \texttt{step-78} TODO

\item \texttt{step-79} TODO
\end{itemize}

@@

The 9.3 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_2_0_and_9_3_0.html}
{around 45 incompatible changes}; see \cite{changes93}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions; and the majority change internal
interfaces that are not usually used in external
applications. That said, the following are worth mentioning since they
may have been more widely used:
\begin{itemize}
  \item The \texttt{hp::DoFHandler} class has been deprecated. The standard
        \texttt{DoFHandler} class is now capable of all $hp$-functionalities.
  \item Consequently, all template arguments \texttt{DoFHandlerType} are obsolete,
        and classes like \texttt{DataOut} or \texttt{SolutionTransfer} for example
        no longer require them. If you rely on these template arguments, an interim
        namespace \texttt{Legacy} has been introduced that provides all affected
        classes with the old interface for the transition phase.
\end{itemize}

@@

The original \dealii{} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{multicols}{2}
  \vspace*{-36pt}
  \begin{itemize}
    \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019};
    \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
    \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007};
    \item For partition-of-unity (PUM) and enrichment methods of the
          finite element space: \cite{Davydov2016};
    \item For matrix-free and fast assembly techniques:
          \cite{KronbichlerKormann2012,KronbichlerKormann2019};
    \item For computations on lower-dimensional manifolds:
          \cite{DeSimoneHeltaiManigrasso2009};
    \item For curved geometry representations and manifolds:
          \cite{HeltaiBangerthKronbichlerMola2019};
          \vfill\null  \columnbreak
    \item For integration with CAD files and tools:
          \cite{HeltaiMola2015};
    \item For boundary element computations:
          \cite{GiulianiMolaHeltai-2018-a};
    \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
          \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b};
    \item For uses of the \texttt{WorkStream} interface:
          \cite{TKB16};
    \item For uses of the \texttt{ParameterAcceptor} concept, the
          \texttt{MeshWorker::ScratchData} base class, and the
          \texttt{ParsedConvergenceTable} class:
          \cite{SartoriGiulianiBardelloni-2018-a};
    \item For uses of the particle functionality in \dealii{}:
          \cite{GLHPB18}.
          \vfill\null
  \end{itemize}
\end{multicols}
